Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-10 Master Thesis/synopsis.tex
Change polynomial variables
Add proof of 4.1

\begin{theorem}[\textit{Hilbert's Irreducibility Theorem}]\label{hilberts-irreducibility-theorem}
Let $E$ be a number field and $f_1,\ldots, f_n$ be irreducible polynomials in $E[X_1,\ldots,X_r,Y_1,\ldots,Y_s]$. Then there exists $a_1,\ldots,a_r\in E$ such that the specializations
\[ f_1(a_1,\ldots,a_r,Y_1,\ldots,Y_s),\ \ldots,\ f_n(a_1,\ldots,a_r,Y_1,\ldots,Y_s) \]
are irreducible in $E[Y_1,\ldots,Y_s]$.
\end{theorem}

@@

\begin{proof}
Let $f: X\to\mathbf{P}^1_E$ be a non-constant morphism and let $f' =  f\circ \pi: X'\to\mathbf{P}^1_E$. By Hilbert's Irreducibility Theorem \ref{hilberts-irreducibility-theorem} there exists a closed $E$-rational point $P$ of $\mathbf{P}^1_E$ which has only one pre-image under $f'$. Consequently, $P$ only has one pre-image under $f$, say $P'$, and $P'$ has only one pre-image under $\pi$.
\end{proof}

@@

\begin{proof}
First we assume that $k/\mathbb{Q}$ is a Galois extension and $X$ has a $k$-rational point $P_0$. We can consider $k'$ as a subfield of $k$ via: 
\[\begin{tikzcd}
k'\ar[r] & k(P'_0)\ar[r, "\varphi_{P_0}"', "\cong"] & k(P_0) \ar[r, "\cong"] & k
\end{tikzcd}\]
For all $P\in X_0$ and $P'=\sigma(P)\in X'_0$ we also get an embedding:
\[ \begin{tikzcd}
k \ar[r] & k(P) \ar[r, "\varphi_P^{-1}"', "\cong"] & k(P')
\end{tikzcd} \]
We see any embedding $k\to \overline{k(P')}$ maps $k$ into $k(P')$ because $k/\mathbb{Q}$ is Galois. Thus we have for all $P'\in X'_0$:
\[ k\otimes_{k'} k(P')\cong \prod_{\operatorname{Hom}_{k'}(k, k(P'))} k(P') \]
in which case $X'_k\to X'$ splits completely. By Lemma \ref{4.1} we get $k\cong k'$, hence $[k:\mathbb{Q}]= [k':\mathbb{Q}]$.

@@

For the uniqueness of $\varphi$ let $\psi: \overline{K}'\to \overline{K}$ be another isomorphism with $\sigma(g) = \psi^{-1} g \psi$. Setting $h = \psi\varphi^{-1}\in G_K$ we observe
\[ h^{-1}gh = \varphi\sigma(g)\varphi^{-1} = \sigma^{-1}\sigma(g) = g \]
for all $g\in G_K$. For $\overline{v}\in\operatorname{Spv}_\text{d}^{(1)}(\overline{K})$ we have therefore $G_{\overline{v}} = h G_{\overline{v}} h^{-1} = G_{h\overline{v}}$, thus $h\in G_{\overline{v}}$ for all $\overline{v}$. Again, by F.K. Schmidt, we have $h=1$ and thus $\varphi = \psi$.
\end{proof}

>>>>

\begin{theorem}[\textit{Hilbert's Irreducibility Theorem}]\label{hilberts-irreducibility-theorem}
Let $E$ be a number field and $f_1,\ldots, f_n$ be irreducible polynomials in $E[X_1,\ldots,X_r,Y_1,\ldots,Y_s]$. Then there exists $e_1,\ldots,e_r\in E$ such that the specializations
\[ f_1(e_1,\ldots,e_r,Y_1,\ldots,Y_s),\ \ldots,\ f_n(e_1,\ldots,e_r,Y_1,\ldots,Y_s) \]
are irreducible in $E[Y_1,\ldots,Y_s]$.
\end{theorem}

@@

\begin{proof}
Let $f: X\to\mathbf{P}^1_E$ be a non-constant morphism and let $f' =  f\circ \pi: X'\to\mathbf{P}^1_E$. This induces a finite field extension $F/E(T)$ of their function fields. Since we're in characteristic $0$, $F/E(T)$ is also separable, and thus we have a primitive element $a\in F$, i.e. $F = E(T,a)$.

Consider the minimal polynomial of $a$ in $E(T)[A]$. Clearing out the denominators, we get an irreducible polynomial $f\in E[T,A]$. By Hilbert's Irreducibility Theorem \ref{hilberts-irreducibility-theorem} we can find an $e\in E$ such that $f(e,A)\in E[A]$ is irreducible.

Now $(T-e)\in\mathbf{A}^1_E$ corresponds to a closed point $P\in \mathbf{P}^1_E$, which in turn corresponds to a valuation in $E(T)$. So the points in the fiber $f'^{-1}(P)$ corresponds to valuations in $F$ above $T-e$. But $T-e$ is inert in $F/E(T)$ since $a\in E[T,a]/(T-e)$ has minimal polynomial $f(e,A)$ over $E[T]/(T-e)$ and $\deg f(e,A) = [F:E(T)]$.

This shows that $P$ has only one pre-image under $f'$. Therefore, it has only one pre-image under $f$, say $P'$, and $P'$ has only one pre-image under $\pi$. This concludes the proof.
\end{proof}

@@

\begin{proof}
First we assume that $k/\mathbb{Q}$ is a Galois extension and $X$ has a $k$-rational point $P_0$. We can consider $k'$ as a subfield of $k$ via: 
\[\begin{tikzcd}
k'\ar[r] & k(P'_0)\ar[r, "\varphi_{P_0}"', "\cong"] & k(P_0) \ar[r, "\cong"] & k
\end{tikzcd}\]
For all $P\in X_0$ and $P'=\sigma(P)\in X'_0$ we also get an embedding:
\[ \begin{tikzcd}
k \ar[r] & k(P) \ar[r, "\varphi_P^{-1}"', "\cong"] & k(P')
\end{tikzcd} \]
We see any embedding $k\to \overline{k(P')}$ maps $k$ into $k(P')$ because $k/\mathbb{Q}$ is Galois. Thus we have for all $P'\in X'_0$:
\[ k(P')\otimes_{k'} k\cong \prod_{\operatorname{Hom}_{k'}(k, k(P'))} k(P') \]
in which case $X'\times_{k'}\operatorname{Spec} (k)\to X'$ splits completely. By Lemma \ref{4.1} we get $k\cong k'$, hence $[k:\mathbb{Q}]= [k':\mathbb{Q}]$.

@@

For the uniqueness of $\varphi$ let $\psi: \overline{K}'\to \overline{K}$ be another isomorphism with $\sigma(g) = \psi^{-1} g \psi$. Setting $h = \psi\varphi^{-1}\in G_K$ we observe
\[ h^{-1}gh = \varphi\sigma(g)\varphi^{-1} = \sigma^{-1}\sigma(g) = g \]
for all $g\in G_K$. For $\overline{v}\in\operatorname{Spv}_\text{d}^{(1)}(\overline{K})$ we have therefore $G_{\overline{v}} = h G_{\overline{v}} h^{-1} = G_{h\overline{v}}$, thus $h\in G_{\overline{v}}$ for all $\overline{v}$. Again, by F.K. Schmidt \ref{theorem-fkschmidt}, we have $h=1$ and thus $\varphi = \psi$.
\end{proof}
